### 2016-10-06 paper/sections/Library.tex

\section{Modular Parsing Library using Delegation}\label{sec:library}

\begin{itemize}
\item Fixpoints library + explaining delegation with some examples
\item Alternative combinator + others
\item Trait Composition to do Language Composition
\end{itemize}