### 2016-10-12 paper/sections/Library/ObjectAlgebras.tex

\subsection{Object Algebras}\label{sec:objectalgebras}