### 2016-11-09 paper/sections/Library/ParsingLibrary.tex
Organize.

\subsection{Fusion of Components/Concepts}

In the previous sections we have talked about open recursion and Object Algebras. These concepts will later be composed together for our
initial goal, which is to build extensible parsers, but surely it cannot live without a parsing library. But we want to argue here that besides
open recursion, which constructs the basis of extensibility in parsing, the other aspects are totally orthogonal, and hence can have their alternatives.
For example, Object Algebras are helpful to build extensible data structures, whereas an alternative with the same function can still be used. And a better
parsing library would introduce a lot of fancy features, nevertheless, it is independent from our key idea.