### 2016-12-26 paper/sections/CaseStudy.tex
Some more comments on Case studies and Implementation

\section{Case Studies}\label{sec:casestudy}

To demonstrate the utility of our modular parsing library, we
implemented parsers of the first 18 calculi from book \textit{Types
  and Programming Languages} (TAPL). TAPL introduces several
calculi from simple to complex, by gradually adding new features to
syntax. \bruno{mention that you compare against an existing
  non-modular implementation.}

\subsection{Overview}\label{subsec:cs-overview}

The calculi in the TAPL book are suitable for our case study for mainly two reasons.
Firstly, the calculi are arguably capture many of the language features 
required in realistic programming languages. Secondly, the evolution of
calculi in the book reveals the advantages of modular representation
of abstract syntax and modular parsing, which is the key functionality
of our library. By decomposing those calculi into smaller components
and reusing the common ones, we can obtain considerably code reuse in
parsing with little performance penalty, as shown later.

\paragraph{Extracting Language Components}
Using the modular parsing techniques discussed before, we extract
reusable \textit{components} from syntax of all the calculi. Each
component, which may contain several syntactical structures, is a
reusable fragment representing a certain feature in the syntax. For
example, \lstinline{VarApp} component represents the variable case and
application case of expressions. In our case study, each component is
represented by a Scala object which has three parts. \lstinline{Alg}
is an object algebra interface for the abstract syntax.
\lstinline{Print} is an object algebra implements that interface
\lstinline{Alg}, for the pretty printing operation. It is not
necessary for parsing, but we include it here for demonstration.
\lstinline{Parser} is the modular parser for this piece of syntax.

\begin{lstlisting}
object VarApp {
  trait Alg[E] {
    def TmVar(x: String): E
    def TmApp(e1: E, e2: E): E
  }
  trait Print extends Alg[String] {
    def TmVar(x: String) = x
    def TmApp(e1: String, e2: String) = "[" + e1 + " " + e2 + "]"
  }
  trait Parser[E, F <: {val pE : PackratParser[E]}] {
    lexical.delimiters += ("(", ")")
    val pVarAppE: Alg[E] => (=> F) => PackratParser[E] = alg => l => {
      val e = l.pE
      ident ^^ alg.TmVar |||
      e ~ e ^^ { case e1 ~ e2 => alg.TmApp(e1, e2) } |||
      "(" ~> e <~ ")"
    }
  }
}
\end{lstlisting}

We use some naming conventions in our code, \lstinline{E} represents
expressions, \lstinline{T} represents types and \lstinline{K}
represents kinds. In this component \lstinline{VarApp} we only have
expressions. The \lstinline{Alg} and \lstinline{Print} are standard
object algebra operations, thus we will focus on the modular parser
\lstinline{Parser}. Each parser carries lexical information and
parsing functions. The lexical information includes reserved words,
delimiters, etc. Parsing functions are similar with we described
before, except that we use record types under some subtyping
constraints here for the explicit self-reference. The record type
contains a parser for expressions \lstinline{pE}, a parser for types
\lstinline{pT} and a parser for kinds \lstinline{pK}, if the
corresponding one is needed when parsing the syntax of this component.
Here the type of self-reference, namely \lstinline{F}, is only
required to have a parser of expressions \lstinline{pE}. Then we can
build the parsing function \lstinline{pVarAppE}, the suffix
\lstinline{E} also denotes that it will parse an expression.

\paragraph{Composing Language Components}
Each calculus could be composed directly from components and other
calculi, if it has those common structures in the syntax. For example,
the calculus \lstinline{Untyped}, representing the famous untyped
lambda calculus, can be constructed from components \lstinline{VarApp}
and untyped lambda abstraction \lstinline{UntypedAbs}.

\begin{lstlisting}
object UntypedAbs {
  trait Alg[E] {
    def TmAbs(x: String, e: E): E
  }
  trait Print extends Alg[String] {
    def TmAbs(x: String, e: String) = "\\" + x + "." + e
  }
  trait Parser[E, F <: {val pE : PackratParser[E]}] {
    val pUntypedAbsE: Alg[E] => (=> F) => PackratParser[E] = {...}
  }
}

object Untyped {
  trait Alg[E] extends UntypedAbs.Alg[E] with VarApp.Alg[E]
  trait Print extends Alg[String] with UntypedAbs.Print with VarApp.Print
  trait Parser[E, L <: {val pE : PackratParser[E]}] extends UntypedAbs.Parser[E, L] with VarApp.Parser[E, L] {
    val pUntypedE = pUntypedAbsE | pVarAppE
  }
}
\end{lstlisting}

When composing the calculus from components, all the fields can be
easily combined by Scala’s \lstinline{extends...with} keywords as
shown in the code. For \lstinline{Parser}, the subtyping constraint
for the explicit self-reference should satisfy all the requirement of
the parsers it extends from. \huang{Talk about | here?}

Finally, we have demo code for every calculus. The parsing result is a
string because we provide the pretty printing algebra to the parser,
and it can be changed as long as the concrete algebra is available. In
this case study we only have pretty printing as the operation, since
it is enough for demonstration.

\begin{lstlisting}
object TestUntyped {
    to be polished
}
\end{lstlisting}

In the next section we will show more examples and talk further about
how we extend and compose these components and calculi.

\subsection{Extensibility}\label{subsec:cs-extensibility}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{resources/depGraph.pdf}
    \caption{Dependency graph of all calculi and components}
    \label{fig:dependency}
\end{figure}

As mentioned before, we decompose all the calculi into reusable
components for modularity. Figure \ref{fig:dependency} shows the
dependency of all the components and calculi in our case study. Gray
boxes are calculi and white boxes are components. An arrow starting
from box A to box B denotes that B includes and thus reuses A.

As shown in the graph, some components such as \lstinline{VarApp} are
created from scratch, while others such as \lstinline{Typed} are
extended from existing components. The \lstinline{Typed} component
extends from \lstinline{VarApp} and has two more syntactical
structures which are typed lambda abstraction and function type (arrow
type). It is the core of many calculi in this case study, and we use
it here as an example for demonstrating the extensibility of
components.

\begin{lstlisting}
object Typed {
  trait Alg[E, T] extends VarApp.Alg[E] {
    def TmAbs(x: String, t: T, e: E): E
    def TyArr(t1: T, t2: T): T
  }
  trait Print extends Alg[String, String] with VarApp.Print {
    def TmAbs(x: String, t: String, e: String) = "\\(" + x + ":" + t + ")." + e
    def TyArr(t1: String, t2: String) = t1 + "->" + t2
  }
  trait Parser[E, T, F <: {val pE : PackratParser[E]; val pT : PackratParser[T]}] extends VarApp.Parser[E, F] {
    ...
    private val pAbsE: Alg[E, T] => (=> F) => PackratParser[E] = {...}
    val pTypedE: Alg[E, T] => (=> F) => PackratParser[E] = pVarAppE | pAbsE
    val pTypedT: Alg[E, T] => (=> F) => PackratParser[T] = alg => l => {
      val t = l.pT
      t ~ ("->" ~> t) ^^ { case t1 ~ t2 => alg.TyArr(t1, t2) } ||| "(" ~> t <~ ")"
    }
  }
}
\end{lstlisting}

Except building a component by extending an existing one, here we have
another extension which is adding a new sort of syntax. It has been
discussed in section \ref{subsec:openrecursion}. In this example,
originally \lstinline{VarApp} has only one sort of syntax which is
expressions, and then \lstinline{Typed} introduces types to the
syntax. We add an extra type parameter \lstinline{T} for it, so that
it is able to be distinguished. The \lstinline{Parser} also requires
the self-reference to have a field \lstinline{pT} for parsing types,
which is represented by the subtyping constraint. With this support,
the \lstinline{pTypedT} function parses arrow types using
\lstinline{pT}. Another point worth mentioning is that we use a
private local function \lstinline{pAbsE} here for parsing the typed
lambda abstraction case, and we can directly combine it with the
inherited \lstinline{pVarApp} to obtain our parsing function of
expressions.

Since calculi and components have the similar signature, each calculus
can also be extended and reused directly. Our last example shows how
to build calculus \lstinline{FullRef} from extending
\lstinline{FullSimple}, another calculus.

\begin{lstlisting}
object FullRef {
  trait Alg[E, T] extends FullSimple.Alg[E, T] with TopBot.Alg[T] with Ref.Alg[E, T]
  trait Print extends Alg[String, String] with FullSimple.Print with TopBot.Print with Ref.Print
  trait Parser[E, T, L <: {val pE : PackratParser[E]; val pT : PackratParser[T]}] extends FullSimple.Parser[E, T, L] with TopBot.Parser[T, L] with Ref.Parser[E, T, L] {
    val pFullRefE = pFullSimpleE | pRefE
    val pFullRefT = pFullSimpleT | pRefT | pTopBotT
  }
}
\end{lstlisting}

Using our modular parsing library, dividing the syntax into small
pieces and bind them with corresponding parsers to form reusable
language component is convenient. From the dependency graph we know
that the common components such as \lstinline{VarApp} are reused in
lots of calculi. Such reuse could shorten the code considerably. We
will show this advantage and exam the possible performance penalty in
a quantitative way in the next section.

\subsection{Comparison}\label{subsec:cs-comparison}

We compared our implementation of the parsers with an implementation
by  Ilya Klyuchnikov, available online. Ilya's
implementation is suitable for comparison, because it is also
written in Scala using the same parser combinator library.
Furthermore, it includes parsers of all the 18 calculi we have, but
written in a non-modular way. Thus it is not able to reuse existing
code of parsing common syntactical features.

Because Ilya's implementation has extra code such as semantics of
those calculi, we removed all irrelevant code and only keep abstract
syntax definition, parser and pretty printer for each calculus, to
ensure a fair comparison.

The comparison is made from two aspects. First, we want to show that
using our modular parsing library, the code can be reused in a
significant way. For this propose, we measured source lines of code
(SLOC) of the two implementations. All blank lines and comment lines
are excluded, and we formatted the code of both implementations so
that the length of each line does not exceed 120 characters. Second,
we are interested to assess the performance penalty caused by modularity. Thus
we compared the execution time, as a major measurement of performance,
between the two implementations. We prepared test files containing
valid expressions in each calculus,\bruno{say something about what
  those expressions are, size, and how they were generated.} then we used the corresponding
parser to parse the file and the pretty printer to print the result.
The execution time excluding reading input file was calculated.

\begin{table}
    \centering
    \begin{tabular}{|l|r|r|r|r|r|r|}
      \hline
        \multirow{2}{*}{\bfseries Calculus Name} & \multicolumn{3}{ c| }{\bfseries SLOC} & \multicolumn{3}{ c| }{\bfseries Time (ms)} \\ \cline{2-7}
        \multicolumn{1}{|c|}{} & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\% & \bfseries Ilya's & \bfseries Ours & \bfseries (+/-)\% \\
      \hline
      \input{resources/CaseStudyTable.tex}
      \hline
    \end{tabular}
    \caption{Comparison of SLOC and execution time}
    \label{tab:comparison}
\end{table}

Table \ref{tab:comparison} shows data of the comparison. The overall
result of SLOC comparison is that 66.3\% of code is reduced using our
library. That is because the code of some common syntax structures
such as variables, lambda abstractions, etc., are reused several times in
the whole project. We can see that in the first two calculi
\lstinline{Arith} and \lstinline{Untyped} we do worse than Ilya's
implementation, because in such two cases we do not reuse any existing
components. But in each of the following 16 calculi, we can compose
our syntax definition, parser and pretty printer from extending
reusable components, resulting considerably code reduction.

\paragraph{Performance}
For the execution time comparison, we built a generator to randomly
generate valid expressions according to the syntax of each calculus.
Each test file corresponds with one calculus and it consists of 500
expressions randomly generated. The size of test files varies from
20KB to 100KB. After testing both implementation, we found that our
version is about \huang{60\%} slower. \bruno{A table showing a summary
of results is missing here. We discussed this before.}
 Part of the reason is that our
library requires more intermediate function calls, and the object
algebra and self-references need to be passed as arguments every
time. However, another important reason is that, because of the
modular setting, we use a different alternative combinator which will
try both branches and match the longest one. \huang{ref to somewhere
  before?} In contrast Ilya uses the ordinary alternative combinator
which tries the second branch only if the first one fails. Although
this difference slows our version, we want to mention that because of
the memoization of Packrat parsers, we are just constant times
slower, the big-O complexity is still the same.