### 2017-01-09 paper/sections/Conclusion.tex
Related work ready for review.

\section{Conclusion}\label{sec:conclusion}

Huang: to write a couple of sentences for a summary.

\noindent Haoyuan: to discuss future work. (some boilerplate can be avoided via metaprogramming.)