### 2017-01-13 paper/sections/Inheritance.tex
Fix refs in Section 2 & 3

\begin{comment}
Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions. We start from Section \ref{subsec:overview-parsing}, which discusses the choice of parsing technique and how it affects modularity of parsers. Section \ref{subsec:overview-problem} demonstrates the goal of extending parsers together with ASTs in a semantic modular way, with both separate compilation and type-safe code reuse. Then we will see traditional parser combinators fail to achieve it because of hard-coded recursive calls. In Section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers. Finally, Section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.\haoyuan{TODO}
\end{comment}

\begin{comment}
It is worth mentioning that the choice of parser combinators will not
affect the other parts of our library. One can choose other parser
combinators like Parsec, in cases that the performance and supporting
of left-recursion are not major concerns. A different library can even build a new
\name with fancy features or higher efficiency.
\end{comment}

\section{Parsing OO ASTs with Multiple Inheritance}\label{sec:inheritance}

Before we address the problem of full modular parsing, we first
address a simpler problem: how to parse Object-Oriented ASTs. To solve
this problem we employ multiple inheritance, which is supported in
Scala via traits.

\subsection{The Mismatch Between OO ASTs and Grammars}
Object-oriented ASTs are ASTs encoded via standard OO techniques,
such as the {\sc Composite} or {\sc Interpreter} patterns~\cite{gamma1995design}. Such ASTs
fix the operations required in the AST, but allow the easy addition of new
language constructs. This is in contrast with ASTs modelled using
functional programming algebraic datatypes or OO style Visitors, where
adding new operations is easy, but adding new language constructs is hard.
In Section~\ref{subsec:packratparsing} we provide a simple example of OO ASTs. The same abstract syntax could be encoded using algebraic datatypes (or alternatively Visitors)
as:

\begin{lstlisting}
data Exp = Lit Int | Add Exp Exp
\end{lstlisting}

Algebraic datatypes and grammars fit well together. Indeed there are
obvious similarities in declaring a grammar, and declaring an
algebraic datatype. It is difficult to modularly add new language
constructs to both grammars and ASTs based on algebraic datatypes. However,
ASTs modelled with algebraic datatypes naturally allow for the easy
addition of operations. So using such ASTs at least preserves one
dimension of extensibility.

Unfortunately, when parsing OO style ASTs, there are difficulties with
the two dimensions of extensibility. On the one hand it is naturally hard to
add new operations to the AST, as usual in the OO style. On the hand
grammars are not modularly extensible, thus the addition of new
language constructs is problematic. Therefore,
it is perhaps not a surprise that parser generators that also generate
AST code (such as ANTLR~\cite{parr2013definitive}), normally generate code based on
visitors. Clearly this seems a better choice since generating code
based on the {\sc Interpreter} or {\sc Composite} pattern would
simply be too inflexible, due to the dual limitation on extensibility.

In the remainder of this section we show a technique for
defining \emph{extensible} parsing code. That is new
language constructs can be modularly added without changing
existing parsing code. The technique employs standard OO
style mechanisms such as subtyping and inheritance, as well as,
Packrat parsing combinators.

\begin{comment}
This section introduces the problem of semantic modular parsing which motivates our work, and an initial solution using only standard inheritance in OO.

\subsection{Modular Parsing Problem}\label{subsec:parsingproblem}
\huang{or "Parsing Extensible ASTs" ?}
The extensibility issues of AST structures and operations that process them, can be illustrated by the famous Expression Problem~\cite{wadler1998expression}. There are two dimensions of extensibility:

\begin{itemize}
\item \textbf{Extension 1}: adding a new data variant (or rather, a new constructor of ASTs).
\item \textbf{Extension 2}: adding a new operation over ASTs.
\end{itemize}

We already have solutions such Object Algebras~\cite{Oliveira2012} for this problem. However, when ASTs evolve, the corresponding parsers which produce such structures should also change accordingly. Our motivation is to build modular parsers, which can be extended together with ASTs. Futhermore, we focus on semantic modularity, that means we expect parsers to be modularly type-checked and separately compiled.

In the rest of Section~\ref{sec:inheritance}, we will introduce an solution for modular parsing with Extension 1 above, using only standard inheritance in OO. Extension 2 will be discussed in Section~\ref{sec:algebrasandparsing}.
\end{comment}

\subsection{Extensible Parsing via Inheritance}

To illustrate how to modularly add new language constructs and the
corresponding parsing code, we continue with the expression language
of literals and additions in Section~\ref{subsec:packratparsing}. We
introduce variables as a new case:

\setlength{\grammarindent}{5em}
\begin{grammar}
<expr> ::= ...
   \alt <ident>
\end{grammar}

It is easy to extend the corresponding OO AST in a modular way:

\lstinputlisting[linerange=29-31]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_SIMPLE_LAM

Since we already have the parser for literals and additions, we would
like to build the new parser by reusing the old one. In Scala we can take advantage of inheritance for such reuse. Specifically,
the new parser can be defined in an enclosing trait that extends
the old one. Here one may quickly come up
with the following attempt, where a new parser is defined for \inlinecode{Var}, then composed
with \inlinecode{pExpr}:

\lstinputlisting[linerange=35-38]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_BAD_ATTEMPT

Unfortunately, this fails to parse some expressions like \inlinecode{"1 + x"}, which are obviously valid in the new grammar.
The reason is that \inlinecode{pAdd} makes two recursive calls to parse sub-expressions, by using \inlinecode{pExpr}, which
covers both cases in the old grammar. But the newly added case \inlinecode{pVar} is not observed by the recursive \inlinecode{pExpr},
hence the parser does not work as expected. It is possible to build the correct parser by replacing the recursive call in \inlinecode{pAdd} with \inlinecode{pVarExpr}.
However, modification on existing code sacrifices separate compilation, and hence breaks semantic modularity.

\subsection{Overriding for Extensibility}\label{subsec:overriding}

It is actually quite simple to let \inlinecode{pExpr} cover the newly extended case without modifying existing code. Method overriding is a standard feature which often comes with inheritance, and it allows us to explain inherited method again, such as \inlinecode{pExpr}. We can build the new parser which correctly parses \inlinecode{"1 + x"} as below.

\lstinputlisting[linerange=42-47]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH

Now \inlinecode{VarExprParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
method overriding in inheritance. When the input \inlinecode{"1 + x"} is fed to the parser \inlinecode{this.pExpr}, it firstly delegates
the work to \inlinecode{super.pExpr}, which parses literals and additions. However, the recursive call \inlinecode{pExpr} in \inlinecode{pAdd}
actually refers to \inlinecode{this.pExpr} again due to dynamic dispatch, and it covers the variable case. Similarly, all recursive calls can be updated to include new extensions if needed.

\subsection{Multiple Extensions and Independent Extensibility}
A nice feature of Scala is its support for a form of multiple inheritance.
Scala has a linearized-style multiple inheritance for traits~\cite{}\huang{what's the proper citation here?}.
This can be very helpful when composing several languages, which
may have been independently developed. Suppose now
we want to compose the parsers for expressions from pre-defined
languages \inlinecode{LanguageA} and \inlinecode{LanguageB} using
alternative. The new parser can be built by inheriting both parsers
at the same time:

\lstinputlisting[linerange=67-73]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=MULTIPLE_INHERITANCE

The \inlinecode{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method \inlinecode{x} in the parent trait \inlinecode{T}. Under multiple inheritance, it can be used to distinguish the method of the same name. Therefore in the new parser, we use \inlinecode{super} to specify which \inlinecode{pExpr} we are referring to.

As a concrete example, we compose the language of integers and additions with a language of boolean values and if-then-else expressions, to build a new language.

\lstinputlisting[linerange=79-95]{../Scala/Parser/src/PaperCode/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE

\begin{comment}
\bruno{You can present the abstract example, as you do
  here, but you should also present a concrete example. You
already have Var, maybe you can add another language extension for
boolean literals. Show the code for the boolean literals parsing as
well as the composition code.}

Interestingly note that, due to the use of multiple inheritance, we
need two different super calls.\bruno{expand here. People may not
be familiar with scala super calls, you have to explain what the
syntax does.}
\end{comment}

Then the new parser, which contains both integer and boolean expressions, is able to parse hybrid expressions such as \inlinecode{"if true then 1 + 2 else 3 + 4"}.

As we demonstrated, inheritance with method overriding is the key technique to obtain semantic modularity.
It enables type-safe code reuse and separate compilation for parsing OO style ASTs.