### 2017-03-21 paper/response.txt
Response draft.

We thank all reviewers for their thoughtful comments. Below we would like to talk about our contributions, and
some questions raised by reviewers.

As we discussed in the paper, almost all previous attempts on extensible/modular parsing stop at syntactical
level of modularity for various reasons. We propose a pattern that nicely achieves semantic modularity and
simplicity for parsing by using standard OO techniques. It’s a notable improvement, since compiling and
generating the parser again during extension every time is obviously a worse practice.

Our framework is indeed based on existing techniques, because a concrete algorithm is required for parsing.
In particular, we use Packrat (PEG) parsing and Object Algebras in our Scala implementation. However, we show
in our case study that they integrate well as expected; the parsing code is elegant and simplified to a great
extent, as boilerplate would explode along with a series of compositions. As discussed in Section 5.4 of the
paper, our approach can also be generalized thus not limited to any specific techniques. What we want to propose
is really a simple, powerful and practical framework, or a flexible platform, instead of a library, to achieve
easy and type-safe composition.

We are sorry that we missed the material of paper [1], as modular parsing and the pattern is not explained
in detail in the paper. Their code online seems to be different with ours as well. Based on parser
combinators, our pattern is more concise and easy to use.

Language components are an important contribution of our work, where we encapsulate a language feature with
three parts: abstract syntax, parser, and operations; everything is modular. With semantically modular parsing,
such components are fully type-safe. Just as SugarJ work and so on, we want languages to be small components,
that can be composed in a flexible way into modular languages again. They are reusable for further composition.
We don't need to rely on a base language to perform extensions as many extensible compilers do.

Rev B:
> However, object algebras solve a problem that is orthogonal to parsing...

We think Object Algebras are not a total orthogonal direction from modular parsing.
Since for parsing, the extension could happen in two cases:
-    When syntax changed, the parser should be extended easily to cooperate with the new syntax.
-    When ASTs changed (a new data variant/new operation added), the old parser should produce
     compatible ASTs without modification.
Regarding the second type of extension, Object Algebras or similar techniques are needed for
modular parsing. Our intuition of using Object Algebras, is not only it encodes multiple syntax
nicely, but also it adds a different dimension to obtain full extensibility of parsing,
supporting separate compilation.

Rev C:
> I would have liked to see it used for more realistic programming language grammars.

We agree that it would be better to show a more realistic language for demonstration. Thanks for your suggestion.

Rev C:
> - Unclear to me if indirect left recursion is supported, which would be needed
> for practical grammars.

Packrat parsing theoretically supports both direct and indirect left recursion [2],
although the implementation in Scala’s standard parser combinator library may be buggy
for some indirect left recursion cases.

Rev C:
> - Does not provide a sufficiently complete solution to language componentization,
> since object algebras are too restricted (only support bottom up computations).

Indeed Object Algebras only support bottom-up traversals, but we have known that some
other researchers have worked out some framework that transforms Object Algebra ASTs
to different ones, which also supports top-down traversals.

Rev E:
> Why is this "type-safety" of the parser so important, after all?  Yes, one may have
> bugs or missing cases in the parser itself, but I would not expect that to be a
> problem, because one should have tests that cover all the grammar anyway, so it is
> little likely that the type-safety of the parser itself would bring much practical benefits.

We think type-safe composition and separate compilation really matter, since it is not a good
idea to modify existing code; we are using OO and want to achieve better code reuse. As mentioned
before, we hope project code to be modular and easily composable, instead of being compiled and
generating code for every round of composition. It is better for debugging.

[1]: Rendel, Tillmann, Jonathan Immanuel Brachthäuser, and Klaus Ostermann. From object algebras to attribute grammars. OOPSLA 2014.
[2]: Alessandro Warth, James R Douglass, and Todd D Millstein. Packrat parsers can support left recursion. PEPM 2008.