### 2017-03-23 paper/response.txt
Response

We thank all reviewers for their thoughtful comments. Below we would
like to talk about our contributions, and some questions raised by
reviewers.

The essential criticism seems to be well summarized by reviewer D:

"Object algebras can be used to modularize tree traversals. Parser
combinators combined with a Packrat algorithm and open recursion allow
modular parser definitions. To people that know both techniques,
object algebras and Packrat parsing, it also appears obvious to
combine the two."

We should also add, on top of the 2 techniques above, there's a third
thing: *good knowledge of some of Scala's more advanced features.*

Is it true that the paper is essentially combining 2 techniques + good
use of Scala to achieve modular parsing, and that people that know
about all of this could possibly recreate this work?

Yes, it is true.

Is that a bad thing?

We are not sure about this. There are at least two arguments
for why the fact that the technique actually ends up being quite
simple is actually a good thing:

1) *It shows that semantically modular parsing can be
seriously considered for practical use*.

Semantic modularization techniques are still far behind syntactic
modularization techniques in terms of proving their
worth/applicability in practice. Without question syntactic modularization
approaches such as those currently used by Language Workbenches,
are currently much more well-accepted in practice. This is in part
because semantic modularization techniques are underdeveloped, and
in part because there is an "established reputation" that semantic
modularization techniques are just simply too hard to use in practice
(and don't scale well).

We believe the simplicity in our work/techniques makes a good
argument for semantic modularization techniques.

2) *We would dare to say that the set of people that knows
Packrat parsing, Object Algebras and Scala well-enough to
know that type-safe modular parsing can be easily done
(before seeing our paper!) is less than 5*

Honestly we didn't know this when we started our work (so we were
not in that set). Although we knew Object Algebras and Scala
well, we were not well-versed with Packrat parsing. We spent quite
a bit of time playing with other parsing combinator libraries, and
getting into trouble before we tried Packrat parsing and solved the
issues we encountered.

Finally, language components and our case study are important
contributions of our work. Language components
encapsulate a language feature with three parts:
abstract syntax, parser, and operations; everything is modular. With
semantically modular parsing, such components are fully
type-safe. Just as SugarJ work and so on, we want languages to be
small components, that can be composed in a flexible way into modular
languages again. They are reusable for further composition.  We don't
need to rely on a base language to perform extensions as many
extensible compilers do. Our case study is also a good benchmark
for other researchers and we believe a good contribution of this work.

Rev B:
> However, object algebras solve a problem that is orthogonal to parsing...

If the goal is to modularize languages (both parsing and semantics),
and have libraries of language components, we don't see how the issues
are orthogonal. The techniques must be combined.

Rev C:
> I would have liked to see it used for more realistic programming language grammars.

We agree that it would be better to show a more realistic language for
demonstration. Thanks for your suggestion.

Rev C:
> - Unclear to me if indirect left recursion is supported, which would be needed
> for practical grammars.

Packrat parsing theoretically supports both direct and indirect left recursion [2],
although the implementation in Scala's standard parser combinator library may be buggy
for some indirect left recursion cases.

Rev C:
> - Does not provide a sufficiently complete solution to language componentization,
> since object algebras are too restricted (only support bottom up computations).

Indeed Object Algebras only support bottom-up traversals, but there's
ongoing work on variants of Object Algebras that remove this
limitation. We think our approach will work just fine with such
variants.

Rev D:

We are sorry that we missed the material of paper [1], as modular
parsing and the pattern is not explained in detail in the paper. Their
code online seems to be different with ours as well. Based on parser
combinators, our pattern seems to be more concise and simpler to use.

Rev E:
> Why is this "type-safety" of the parser so important, after all?  Yes, one may have
> bugs or missing cases in the parser itself, but I would not expect that to be a
> problem, because one should have tests that cover all the grammar anyway, so it is
> little likely that the type-safety of the parser itself would bring much practical benefits.

We think type-safe composition and separate compilation really matter,
since it is not a good idea to modify existing code; we are using OO
and want to achieve better code reuse. As mentioned before, we hope
project code to be modular and easily composable, instead of being
compiled and generating code for every round of composition. It is
better for debugging.

[1]: Rendel, Tillmann, Jonathan Immanuel Brachthäuser, and Klaus Ostermann. From object algebras to attribute grammars. OOPSLA 2014.
[2]: Alessandro Warth, James R Douglass, and Todd D Millstein. Packrat parsers can support left recursion. PEPM 2008.