### 2017-06-07 SLE_17/sections/Inheritance.tex
12 pages.

\begin{comment}
Basically, \name consists of four parts: underlying parsing technique, delegation mechanism encoded by open recursion, Object Algebras, and glue code of new combinators and utility functions. We start from Section \ref{subsec:overview-parsing}, which discusses the choice of parsing technique and how it affects modularity of parsers. Section \ref{subsec:overview-problem} demonstrates the goal of extending parsers together with ASTs in a semantic modular way, with both separate compilation and type-safe code reuse. Then we will see traditional parser combinators fail to achieve it because of hard-coded recursive calls. In Section \ref{subsec:overview-delegation}, we show how delegation can solve this problem and allow us to build extensible parsers. Finally, Section \ref{subsec:overview-oa} gives examples of using Object Algebras for more extensibility, including extension of operations and parsing multiple sorts of syntax.\haoyuan{TODO}
\end{comment}

\begin{comment}
It is worth mentioning that the choice of parser combinators will not
affect the other parts of our library. One can choose other parser
combinators like Parsec, in cases that the performance and supporting
of left-recursion are not major concerns. A different library can even build a new
\name with fancy features or higher efficiency.
\end{comment}

\section{\hspace{-10pt}OO AST Parsing with Multiple Inheritance}\label{sec:inheritance}

Before we address the problem of full modular parsing, we first
address a simpler problem: how to parse Object-Oriented ASTs. To solve
this problem we employ multiple inheritance, which is supported in
Scala via traits.

\begin{comment}

\subsection{The Mismatch Between OO ASTs and Grammars}\label{subsec:mismatch}
Object-oriented ASTs are ASTs encoded via standard OO techniques,
such as the {\sc Composite} or {\sc Interpreter} patterns~\cite{gamma1995design}. Such ASTs
fix the operations required in the AST, but allow the easy addition of new
language constructs. We have seen an example of OO ASTs in Section~\ref{subsec:packratparsing}.
In contrast, ASTs modelled using functional programming algebraic datatypes (or alternatively OO style Visitors) are as follows:

\begin{lstlisting}
data Exp = Lit Int | Add Exp Exp
\end{lstlisting}

Algebraic datatypes and grammars fit well together. Indeed there are
obvious similarities in declaring a grammar, and declaring an
algebraic datatype. Unlike OO ASTs, they naturally allow for the easy
addition of operations, whereas modularly adding new constructs to grammars
or ASTs becomes hard. Using this style of ASTs again preserves one dimension of extensibility.

Unfortunately, when parsing OO style ASTs, there are difficulties with
the two dimensions of extensibility. On the one hand it is naturally hard to
add new operations to the AST, as usual in the OO
style~\cite{wadler1998expression}. On the other hand
grammars are not modularly extensible, thus the addition of new
language constructs is problematic. Therefore,
it is perhaps not a surprise that parser generators also generate
AST code (such as ANTLR~\cite{antlr1995}), normally the generated code is based on
visitors. Clearly this seems a better choice since generating code
based on the {\sc Interpreter} or {\sc Composite} pattern would
simply be too inflexible, due to the dual limitation on extensibility.

In the remainder of this section we show a technique for
defining \emph{extensible} parsing code. That is new
language constructs can be modularly added without changing
existing parsing code. The technique employs standard OO
style mechanisms such as subtyping and inheritance, as well as,
Packrat parser combinators, and enables a better integration between
parsing and OO ASTs.

\end{comment}

\begin{comment}
This section introduces the problem of semantic modular parsing which motivates our work, and an initial solution using only standard inheritance in OO.

\subsection{Modular Parsing Problem}\label{subsec:parsingproblem}
\huang{or "Parsing Extensible ASTs" ?}
The extensibility issues of AST structures and operations that process them, can be illustrated by the famous Expression Problem~\cite{wadler1998expression}. There are two dimensions of extensibility:

\begin{itemize}
\item \textbf{Extension 1}: adding a new data variant (or rather, a new constructor of ASTs).
\item \textbf{Extension 2}: adding a new operation over ASTs.
\end{itemize}

We already have solutions such Object Algebras~\cite{Oliveira:2012} for this problem. However, when ASTs evolve, the corresponding parsers which produce such structures should also change accordingly. Our motivation is to build modular parsers, which can be extended together with ASTs. Futhermore, we focus on semantic modularity, that means we expect parsers to be modularly type-checked and separately compiled.

In the rest of Section~\ref{sec:inheritance}, we will introduce an solution for modular parsing with Extension 1 above, using only standard inheritance in OO. Extension 2 will be discussed in Section~\ref{sec:algebrasandparsing}.
\end{comment}

Part of the modular parsing problem is how to obtain an extensible parser. It
is natural to make use of OO ASTs because adding new data constructs is cheap for
them. Hence we have used OO traits and inheritance to represent the AST in the last
section. Furthermore, we would like to write extensible parsing code on extensions of
a grammar. That is to say, new extensions would not require modifying the existing code,
and we can even reuse the old code.

To illustrate such extensibility, we continue with the old
example, and introduce variables as a new case.
It is easy to extend the corresponding OO AST together with its parser in a modular way:

\begin{lstlisting}
class Var(x: String) extends Expr {
  def print = x
}
\end{lstlisting}
Here one may quickly define a new parser \lstinline|pVar| in \lstinline|AParser| for variables, and parse new expressions
with \lstinline{"pExpr ||| pVar"}.
Unfortunately, even \inlinecode{"1 + x"} cannot be parsed, which is obviously valid in the new grammar.
The reason is that \inlinecode{pAdd} makes two recursive calls to \lstinline|pExpr| for parsing sub-expressions, whereas the newly added \inlinecode{pVar} is not observed, unless we replace all the occurrences of \lstinline|pExpr| with \lstinline{"pExpr ||| pVar"}.
Yet modifying existing code breaks semantic modularity.

\vspace{-.05in}
\paragraph{Overriding for Extensibility}
It is actually quite simple to let \inlinecode{pExpr} cover the newly
extended case without modifying existing code. Method overriding is a
standard feature which often comes with inheritance, and it allows us
to redefine an inherited method, such as \inlinecode{pExpr}. We can
build the new parser which correctly parses \inlinecode{"1 + x"}
through overriding:

\lstinputlisting[linerange=20-26]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=INHERITANCE_APPROACH

Now \inlinecode{VarParser} successfully represents the parser for the extended language, because Scala uses dynamic dispatch for
method overriding in inheritance. When the input \inlinecode{"1 + x"} is fed to the parser \inlinecode{this.pExpr}, it firstly delegates
the work to \inlinecode{super.pExpr}, which parses literals and additions. However, the recursive call \inlinecode{pExpr} in \inlinecode{pAdd}
actually refers to \inlinecode{this.pExpr} again due to dynamic dispatch, and it covers the variable case. Similarly, all recursive calls can be updated to include new extensions if needed.

\vspace{-.05in}
\paragraph{Independent Extensibility}
A nice feature of Scala is its support for the linearized-style multiple
inheritance on traits~\cite{odersky2004overview}. This can be very helpful when
composing several languages, and to achieve independent
extensibility~\cite{odersky2005independently}. Suppose now we want to compose the parsers
for expressions from pre-defined languages \lstinline|LanguageA| and \lstinline|LanguageB| using alternative.
The new parser can be built by inheriting both parsers at the same time:
\begin{lstlisting}
trait LanguageA {...}
trait LanguageB {...}
trait LanguageC extends LanguageA with LanguageB {
  override def pExpr = super[LanguageA].pExpr ||| super[LanguageB].pExpr
}
\end{lstlisting}

The \lstinline{super[T].x} syntax in Scala, so called \textit{static super reference}, refers to the type or method
\inlinecode{x} in the parent trait \inlinecode{T}. Under multiple inheritance, it can be used to distinguish
the methods of the same name. Therefore in the new parser, we use \lstinline{super} to specify which
\lstinline{pExpr} we are referring to.

%As a concrete example, we compose the language of integers and additions with a language of boolean values and if-then-else expressions, to build a new language.

%\lstinputlisting[linerange=56-68]{code/src/papercode/Sec3Inheritance/Code1.scala}% APPLY:linerange=BOOL_EXAMPLE

\begin{comment}
\bruno{You can present the abstract example, as you do
  here, but you should also present a concrete example. You
already have Var, maybe you can add another language extension for
boolean literals. Show the code for the boolean literals parsing as
well as the composition code.}

Interestingly note that, due to the use of multiple inheritance, we
need two different super calls.\bruno{expand here. People may not
be familiar with scala super calls, you have to explain what the
syntax does.}
\end{comment}

%Then the new parser, which contains both integer and boolean
%expressions, is able to parse hybrid expressions such as
%\inlinecode{"if true then 1 + 2 else 3 + 4"}.

\vspace{-.05in}
\paragraph{Conflicts and/or Ambiguity} In a modular setting, conflicts and
ambiguity could be introduced to the grammar. In that case, the help parser combinators
can offer is quite restricted. Yet users can override those problematic methods to resolve
such conflicts, and rely on dynamic dispatch. We will discuss it in Section~\ref{subsec:overriding-rules}.

As demonstrated, inheritance with method overriding is the key
technique to obtain semantic modularity.  It enables type-safe code
reuse and separate compilation for parsing OO style ASTs.