### 2017-09-04 SLE_17/sections/RelatedWork.tex

\section{Related Work}\label{sec:relatedwork}

%- extensible parsing, language workbenches: rats, noa (this one already uses OA), modular semantic actions, (syntactic modularity, no separate compilation, modular type-checking)
%(read more papers, see if they talk about this issue, some potential solutions)
%(attribute grammars?)
%
%- parser combinators for type-checking, previous work has not shown how to support modularity (ASTs); left-recursion and back-tracking in related techniques
%
%- modularity: object algebras, dtc and mrm (problem with parsing, is there any related work? (PB: a paper on unfolds: build the AST))
%
%(parsing in Javascript: using delegation, does it support modular AST)
%
%noa, shy: shy: only override some interesting cases (transformation is tedious)
%bruijn indices: parsing + transformation

Our work touches upon several topics including extensible parsing,
parser combinators and extensibility techniques. 
%However, as far as we
%know there's no work that discusses how to do statically type-safe and
%separately compilable modular parsing.

\begin{comment}
There has been a
great amount of related papers on those topics. Some
inspired us of this paper and encourage us for more exploration. This
section will try to lead a discussion on what difference we have made.
\end{comment}

\paragraph{Syntactically Extensible Parsing}
Extensible parser generators~\cite{antlr1995,Grimm2006,Gouseti2014,Viera2012,Warth2016,schwerdfeger09}
are a mainstream area of modular syntax and parsing. They allow users to write
modular grammars, where new
non-terminals and production rules can be introduced, some can even
override existing rules in the old grammar modules. For instance,
\textit{Rats!}~\cite{Grimm2006} constructs its own module system for
the collection of grammars, while NOA~\cite{Gouseti2014} uses Java
annotation to collect all information before producing an ANTLR~\cite{antlr1995} grammar
and the parsing code. Those parser
generators focus on the \textit{syntactic extensibility} of grammars:
they rely on whole compilation to generate a global parser, even if
there is only a slight modification in the grammar. Some of those
parser generators may statically check the correctness and unambiguity
of grammars. In contrast, because our approach is based on parser
combinators, there is no support for ambiguity checking.  However, as
far as we are aware, no extensible parser generators support separate
compilation or modular type-checking. It is worth mentioning that in~\cite{Viera2012}, users can define
grammar fragments as typed Haskell values, and combine them on the fly. Later they are processed
by a typed parser generator. Nevertheless this requires
a lot of advanced language features, making client complex. Our
approach is simple and a straightforward use of OO programming, 
and makes parsing code directly reusable.

Macro systems like the C preprocessor, C++ templates and
Racket~\cite{Tobin-Hochstadt2011}, and other meta-programming
techniques are a similar area aiming at syntactic extensibility.
SugarJ~\cite{Erdweg2011} conveniently introduces syntactic sugar for
Java using library imports. Composition of syntactic sugar is easy for
users, but it requires many rounds of parsing and adaption, hence
significantly affects the efficiency of compilation. Since the
implementation was based on SDF~\cite{Heering1989} and
Stratego~\cite{Visser2001}, it does not support separate
compilation. Racket adopts a macro system for library-based language
extensibility~\cite{Tobin-Hochstadt2011}. It uses
attributed ASTs for contextual
information, and extensions can be integrated in a modular
way. However such modularity is not flexible enough for language
unification, as the syntax is only built from extensions.
%Moreover,
%existing macros cannot be further changed after
%definition. \haoyuan{correct?}
Extensible
compilers like JastAdd~\cite{Ekman2007} and
Polyglot~\cite{Nystrom2003} also support extensible parsing, but it
is mostly done using parser generators. They focus on the
extensions to a host language. Those techniques are short of type safety in a modular
setting as well.

Besides, type-specific languages~\cite{omar14} do support
separate compilation and type safety. The modularity of parsing is incorporated in its type
system, which allows users to introduce new syntax for parsing. Whereas in our approach, parsing 
strategy is an orthogonal issue since we rely on an external parsing library.
Our approach does not rely on an elaborated semantics extension, but simply uses a pattern
that is applicable in mainstream OO languages, where the type system helps ensure
type safety.

%\bruno{Have you read
%Languages as Libraries? That is probably an important reference, which
%I think we should cite.}
%\haoyuan{what about
%  metafront? it is a macro system but does it have type safety?}
%\haoyuan{Extensible syntax with lexical scoping?} \haoyuan{"Extensible
%  syntax" proposes a system for extensible syntax, where users write
%  EDSLs in their language with concrete syntax. Users can write rules
%  for type-based disambiguation. But separate compilation is again not
%  mentioned. Shall we mention that thesis?} \haoyuan{attribute
%  grammars?}

\paragraph{Extensible Parsing Algorithms}
\textit{Parse table composition}~\cite{bravenboer2008parse, schwerdfeger2010}
is an approach where grammars are compiled to
modular parse tables. Those parse tables are expressed as DFAs
or NFAs, and later they can be composed by an algorithm, to provide
separate compilation for parsing. The generation of parse tables can
be quite expensive in terms of performance. The approach
is quite different from ours, since it uses parse
 tables, whereas we use parser combinators.
Our approach supports both
\emph{separate compilation} as well as \emph{modular
  type-checking}, and is applied in common OO languages. Moreover, the extensibility of parsing is further
available at language composition.

%\haoyuan{we do not have explicit
%  correspondence/relationship between abstract syntax and the parser.}

\paragraph{Parser Combinators} Parser combinators have become more and more
popular since~\cite{burge1975,Wadler1985}. Many parsing libraries produce recursive descent
parsers by introducing functional monadic
parser combinators~\cite{nott237}. Parsec~\cite{Leijen2001} is
perhaps the most popular parser combinator library in this line.
It is widely used in Haskell (with various ``clones'' in other languages)
for context-sensitive grammars with infinite lookahead. Nevertheless,
Parsec users suffer from manual left-recursion elimination,
high cost for backtracking and longest match composition issues,
as we discussed in Section~\ref{subsec:challenges}. Those limitations make Parsec
(and similar parsing techniques) inadequate for modular parsing.

Some recent work on parser
combinators~\cite{Ford2002,Might2011,Frost2008} proposed a series of
novel parsing techniques that address the issue of
left-recursion. We chose Packrat parsing due to its simplicity in Scala,
but in general there are alternatives to it.

\paragraph{Extensibility} Various design patterns~\cite{gamma1995design} in multiple
languages, have been proposed over the years to address extensibility
problems, such as the Expression Problem~\cite{wadler1998expression}.
The famous ``Datatypes \`a
la Carte'' (DTC)~\cite{swierstra2008data} approach represents modular ASTs using co-products
of every two functors. Several variants of DTC have been later proposed~\cite{Bahr2011,Bahr2014,Oliveira2015}.
All of that work essentially covers how to traverse and
consume extensible ASTs. However they do not
address the problem of \emph{modularly
parsing extensible ASTs}. Only in Bahr's~\cite{Bahr2011} work \emph{unfolds} is briefly mentioned,
yet it does not cover parsing.

There are also many design patterns in OO languages that achieve
type-safe extensibility~\cite{torgersen2004expression,odersky2005independently,oliveira2009modular,Oliveira:2012,wang2016expression}. We chose Object Algebras~\cite{Oliveira:2012} because the pattern is
relatively lightweight and makes good use of existing OO features,
such as inheritance, generics and subtyping. As seen throughout the paper,
the parsing code is concise and expressive using Object Algebras.
On the other hand we are unware of any work on OOP
that has covered how to do modular parsing for extensible ASTs.

It is worth mentioning that Scala case classes~\cite{emir2007matching} provide a near
solution to the Expression Problem. Case classes can be
modularly added, but they do not enforce
exhaustiveness of pattern matching for extensible operations. In other
words run-time pattern matching errors can happen when writing extensible code with case classes. So that full static type-safety is not ensured. Nevertheless case classes are a pragmatic approach, which is
widely used in practice. Therefore, modular parsing techniques may be
of value for extensible code using case classes.  The approach we
presented in Section~\ref{sec:inheritance} can readily be adapted to case classes:
all that the users need to do is to use case classes, instead of
standard OO classes in their code.

\begin{comment}
Moreover, in~\cite{Oliveira:2012} the authors have discussed the
composition of algebras. In our parsing approach, a parser consumes an
algebra, which is delegated to return the results, during its process
of parsing. Having a set of algebras, it requires multiple parsing
with several times of invocation, which leads to redundant work.
Instead, algebras are supposed to be composed into one before the
invocation of the parser. Bahr et al. lead a similar discussion
in~\cite{Bahr2011}, where queries (or \textit{catamorphisms}) and
transformations (or \textit{homomorphisms}) are composable. They have
also mentioned the dual process of folds, namely
\textit{anamorphisms}. It is potentially related to our work, as
parsing is a representative kind of unfolds, whereas they only
discussed the composition of a cv-coalgebra and a term homomorphism,
which differs from modular parsing.
\end{comment}

\begin{comment}
Another interesting observation from Section~\ref{sec:algebrasandparsing}, is that parsers with
Object Algebras generate ``implicit objects'', which are actually functions of
type \lstinline{Alg[E]} \lstinline{=>} \lstinline{E} for generic \lstinline{E}.
Patterns of operations on such functions are captured by the original paper~\cite{Oliveira:2012}, called
\textit{queries} and \textit{transformations}, together with the composition of algebras.
They are quite useful, for instance, we have two operations (like the aforementioned \lstinline{Print} and \lstinline{Eval}) to be fed to the parser. If we feed them separately, it is tedious to parse the same input twice. Instead we can compose them in advance before the feeding. On the other hand, one would think this pattern is too limited to use against objects that traditional parsers generate. For example, there are a lot of transformations (or desugarings) on intermediate ASTs in the design of a compiler. Transformation algebras are helpful in that case. They can even be composed in a linear pipeline-style before a query algebra is delegated. The Shy framework~\cite{Zhang2015} has captured those patterns, and it generates templates for them, so that users only need to override a few interesting cases, conveniently. The idea of overriding existing cases was also
proposed in MRM~\cite{Oliveira2015}. They are potentially useful to our approach. Regarding overriding, our approach allows existing parsing code to be overridden during inheritance.
\end{comment}